# Introduction to ggplot2

```{r}
library(PKPDmisc)
library(tidyverse)
library(ggplot2)
```

## XY Plots

1) For concentration-time plots filter only OCC 1 from dapa_IV_oral (in PKPDdatasets) for subsequent plotting

```{r}
data1 <- dapa_IV_oral %>% filter(OCC == 1)
data1 <- data1 %>% mutate(GENDER = ifelse(GENDER == 0, "Male","Female"))
```

2) Basic Concentration-Time plot (point and lines)

```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot")

```



3) make points/lines bigger
```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS), size = 3) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID), size = 2) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot")

```


4) add log transformation to y axis

```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID)) + 
    xlab("Time") + ylab("Concentration") + scale_y_log10() + ggtitle("Concentration-Time plot")

```

5) Add color by gender

```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS, color = GENDER)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID, color = GENDER)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot")

```

BONUS: rename labels legend

```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS, color = GENDER)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID, color = GENDER)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot")
```
BONUS: move legend to top right
```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS, color = GENDER)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID, color = GENDER)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot") + theme(legend.justification = "top")
```

7) rename facet strips by name
```{r}
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS, color = GENDER)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID, color = GENDER)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot (faceted by ID)") +    theme(legend.justification = "top") + facet_wrap(~ ID, nrow = 2)
ggplot(data = data1) + geom_point(mapping = aes(x = TIME, y = COBS, color = GENDER)) +
  geom_line(mapping = aes(x = TIME, y = COBS, group = ID, color = GENDER)) + 
    xlab("Time") + ylab("Concentration") + ggtitle("Concentration-Time plot") +    theme(legend.justification = "top") + facet_wrap(~ GENDER, nrow = 2)

```


8) rename axes



## Boxplots and Histograms

* Histogram(s) of demographics

```{r}
# distinct is a dplyr verb - easiest way to subset data by the 
# first row in some combination.
sid_data <- data1 %>% 
    distinct(ID, .keep_all = TRUE) # single row per id data

ggplot(sid_data, aes(AGE)) + geom_histogram() + 
    xlab("Age") + ylab("Number of subjects") + ggtitle("") 

```

* add vertical line for median value
```{r}

ggplot(sid_data, aes(AGE)) + geom_histogram() + 
    xlab("Age") + ylab("Number of subjects") + ggtitle("") + geom_vline(aes(xintercept = median(AGE)), size = 2)

```
* put the numeric value of the vertical line as text next to the line on the plot
```{r}

ggplot(sid_data, aes(AGE)) + geom_histogram() + 
    xlab("Age") + ylab("Number of subjects") + ggtitle("") + geom_vline(aes(xintercept =  median(AGE)), size = 2, linetype = "dashed") + geom_text(data = sid_data, aes(x = median(AGE), y = 0, label = median(AGE)), size = 4, vjust = -1.5)

```

```{r}
devtools::session_info()
```

